### 2015-06-22 Talks/Spurgin_Rebellion_2015.Rmd

---
output:
  beamer_presentation:
    fig_caption: false
    colortheme: 'seahorse'
---

```{r setwd, include = F}

#Set working directory up one level
setwd('..')

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

#Set some options

opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE)

```

## Good Years, Bad Years and Telomeres in a Wild Bird Population

\begin{center}
  \includegraphics{Pictures/SWfeedingcomp2.png} \\~\\

Lewis Spurgin \\~\\

March 2015 \\~\\

  \includegraphics{Pictures/UEANERCwide.png}

\end{center}

## The benefits of a good start

\begin{center}
  \includegraphics{Pictures/Bullingdon.png}
\end{center}

## The benefits of a good start

\begin{center}
  \includegraphics{Pictures/Slytherin.png}
\end{center}

## A good/bad start doesn't affect everyone equally

\begin{center}
  \includegraphics{Pictures/powers.png}
\end{center}

## Silver spoons in ecology

\begin{center}
  \includegraphics{Pictures/DeerFig.png}
\end{center}

## Silver spoons in ecology

\begin{center}
  \includegraphics{Pictures/GreatTitFig.png}
\end{center}

## Can we quantify how badly an individual suffers from a bad start in life?

\begin{center}
  \includegraphics{Pictures/BirdOfParadise.png}
\end{center}

## Telomeres - (TTAGGG)~*n*~

\begin{center}
  \includegraphics{Pictures/Telomeres.png}
\end{center}

## Telomeres and oxidative stress

\begin{center}
  \includegraphics{Pictures/Oxidants.png}
\end{center}

## Telomeres - biomarkers of cost

\begin{center}
  \includegraphics{Pictures/Google.png}
\end{center}

## Telomeres and silver spoons

\begin{center}
  \includegraphics{Pictures/ShagFig.png}
\end{center}

## Telomeres and silver spoons

\begin{center}
  \includegraphics{Pictures/HeidingerFig.png}
\end{center}

## Seychelles warbler (*Acrocephalus sechellensis*)

\begin{center}
  \includegraphics{Pictures/CousinAndWarbler.png}
\end{center}

## Seychelles warbler (*Acrocephalus sechellensis*)

\begin{center}
  \includegraphics{Pictures/SWnest.png}
\end{center}

## It rains in the tropics

\begin{center}
  \includegraphics{Pictures/Rain.png}
\end{center}

## Food availiability varies over time

```{r food availability over time,fig.width=4,fig.height=4}
temp <- subset(FlSAall,LayYear!='1996')

temp$LayYear <- as.numeric(paste(temp$LayYear))
Fig1A <- ggplot(temp,
       aes(x = LayYear,
           y = Insect)) +
  xlab('Year') + 
  ylab('Insect availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14),
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_ribbon(aes(ymax=Insect+InsectSE,ymin=Insect-InsectSE),alpha=0.2) +
  geom_path()

Fig1A

```

## Food availiability in early life doesn't predict lifespan in the Seychelles warbler

```{r Lifespan and food availability,fig.width=4,fig.height=4}

temp <- FlSAall
temp$Insect <- factor(temp$Insect)

ddFig1A <- ddply(temp,
                 .(Insect),
                 summarize,
                 TLKBmean = mean(RemainingLife),
                 TLKBse = 1.96*se(RemainingLife))
ddFig1A <- ddFig1A[complete.cases(ddFig1A),]
ddFig1A$Insect <- as.numeric(paste(ddFig1A$Insect))

Fig1A <- ggplot(ddFig1A,
       aes(x = Insect,
           y = TLKBmean)) +
  ylab('Lifespan') + 
  xlab('Food availability in early life') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +

  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse))

Fig1A

```

## Predictions

**Telomeres link early life conditions to late life survival**

- Juvenile telomere length varies over years

- Telomere length in early life is related  to food availability

- Individuals who suffer greatest telomere shortening in early life have reduced late life survival

\begin{center}
  \includegraphics{Pictures/BirdOfParadiseSmall.png}
\end{center}

## Measuring warbler telomeres

\begin{center}
  \includegraphics{Pictures/Emma.png}
\end{center}

## Juvenile telomere length varies over years

```{r Telomere length over time,fig.width=4,fig.height=4}

temp <- FlSA

ddFigS2 <- ddply(temp,
                 .(LayYear),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS2 <- ddFigS2[complete.cases(ddFigS2),]
ddFigS2$LayYear <- as.numeric(paste(ddFigS2$LayYear))

FigS2 <- ggplot(ddFigS2,
       aes(x = LayYear,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  geom_line() +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS2

```

## Telomere length in early life is related  to food availability

```{r Telomeres and food availability,fig.width=4,fig.height=4}

temp <- FlSAall
temp$Insect <- factor(temp$Insect)

ddFig1A <- ddply(temp,
                 .(Insect),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = 1.96*se(TLKB))
ddFig1A <- ddFig1A[complete.cases(ddFig1A),]
ddFig1A$Insect <- as.numeric(paste(ddFig1A$Insect))

Fig1A <- ggplot(ddFig1A,
       aes(x = Insect,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('Food availability in early life') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  stat_smooth(data=FlSAall,aes(x=Insect,y=TLKB),method=lm,col=grey(0.7)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse))

Fig1A

```

## Individuals who suffer greatest telomere shortening in early life have reduced late life survival

```{r Telomeres and survival,fig.width=5.5,fig.height=4}

Fig2 <- ggplot(subset(subset(FlSAall,RemainingLife>1),RemainingLife<11),
                aes(y = RemainingLife,
                x = TLKB,
                col=YQ)) +
  theme_classic() +
  geom_point(alpha=0.3) +
  xlab('Telomere length (kb)') + 
  ylab('Remaining life') + 
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16,vjust = 1), 
        legend.text = element_text(size = 13),
        legend.title = element_blank()) +
  scale_color_manual(values=c('#d95f02','#1b9e77'))+
  scale_fill_manual(values=c('#d95f02','#1b9e77'))+
  stat_smooth(method=glm,family=poisson,lwd=1.2,aes(fill=YQ,lty=YQ),fullrange=T)

Fig2

```

## Predictions

**Telomeres link early life conditions to late life survival**

- Juvenile telomere length varies over years

- Telomere length in early life is related  to food availability

- Individuals who suffer greatest telomere shortening in early life have reduced late life survival

\begin{center}
  \includegraphics{Pictures/BirdOfParadiseSmall.png}
\end{center}

## Predictions

**Telomeres link early life conditions to late life survival**

- Juvenile telomere length varies over years **YES**

- Telomere length in early life is related  to food availability

- Individuals who suffer greatest telomere shortening in early life have reduced late life survival

\begin{center}
  \includegraphics{Pictures/BirdOfParadiseSmall.png}
\end{center}

## Predictions

**Telomeres link early life conditions to late life survival**

- Juvenile telomere length varies over years **YES**

- Telomere length in early life is related  to food availability **YES**

- Individuals who suffer greatest telomere shortening in early life have reduced late life survival

\begin{center}
  \includegraphics{Pictures/BirdOfParadiseSmall.png}
\end{center}

## Predictions

**Telomeres link early life conditions to late life survival**

- Juvenile telomere length varies over years **YES**

- Telomere length in early life is related to food availability **YES**

- Individuals who suffer greatest telomere shortening in early life have reduced late life survival **YES**

\begin{center}
  \includegraphics{Pictures/BirdOfParadiseSmall.png}
\end{center}

## Thank you!

\begin{center}
  \includegraphics{Pictures/Acks.png}
\end{center}