### 2015-09-01 Manuscript/SW_early_life_telomeres_tables_and_figures v9.Rmd

---
output: word_document
---
```{r setwd, include = F}

#Set working directory up one level
setwd('..')

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = getwd())
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.show='asis',
               dev='png')

```

**Figure Legends**

**Figure 3** Telomere length and age in juvenile Seychelles warblers. Points and error bars represent mean and 95% confidence intervals, respectively.

**Figure 2** Factors affecting telomere length in Seychelles warbler chicks. **A** Model averaged estimates and 95% confidence intervals (points and error bars, respectively) for all explanatory terms used in a linear model with juvenile telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality). **B** Presence/absence of helpers and **C** body mass in relation to raw telomere length. Points and error bars in **B** are mean $\pm$ 95% confidence interval telomere lengths for each group, and the line and shaded areas from **C** represent fitted values and 95% confidence limits from a linear regression.

**Figure 3** Factors affecting telomere length in juvenile Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals (points and error bars, respectively) for all explanatory terms used in a linear model with juvenile telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality). **B** Raw telomere length in relation to annual food availability Points and error bars in **B** are mean $\pm$ 95% confidence interval telomere lengths for each year (means are used for plotting purposes only), and the line and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 4** Juvenile telomere length and survival in chick (**A,B**) and juvenile (**C,D**) Seychelles warblers. **A,C** Model averaged estimates and 95% confidence intervals (points and error bars, respectively) for all explanatory terms used in a parametric survival model (see main text for details). Food av = annual food availability, Tel. length = telomere length. **B,D** Kaplan-Meier curves showing the relationship between telomere length and survival. Telomere length was modelled as a covariate, but is binned into groups here for visualisation purposes (long and short = greater than or less than median telomere length, repectively).

PAGEBREAK

**Figure 1**

```{r Figure 1,fig.width=5,fig.height=4}

temp <- dd

ddFig1 <- ddply(temp,
                 .(AgemonthF),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB)*1.96)

Fig1 <- ggplot(ddFig1,
       aes(x = AgemonthF,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('Age (months)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse))

Fig1

```

PAGEBREAK

**Figure 2**

```{r Figure 2,fig.width=8,fig.height=9}

###########
#Fig 2A

chickall <- subset(chickall,BodyMass>5)
#Run the model
lm2 <- lmer(TLKB ~ TQspace + Insect + BodyMass+ Tarsus + NonHelper + Sex + Helper + (1|LayYear),
            data = chickall,REML=FALSE)

lm2 <- standardize(lm2)

#Model averaging
lm3 <- dredge(lm2)
lmavg <- model.avg(lm3,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')

table2 <- data.frame(table2,myx=row.names(table2))

Fig2A <- forestplot(table2) +
  annotate('text',x=0.75,y=4,label='A')

######################################
temp <- chickall
temp$Helper <- ifelse(temp$Helper>0,'Present','Absent')

ddFig2B <- ddply(temp,
                 .(Helper),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = 1.96*se(TLKB))

Fig2B <- ggplot(ddFig2B,
       aes(x = Helper,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('Helpers') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  scale_colour_manual(values = wes_palette("Royal1")) +
  annotate('text',x=0.65,y=9,label='B')

Fig2C <- ggplot(chickall,
       aes(x = BodyMass,
           y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Body mass (g)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point() +
  stat_smooth(method = 'lm',col = 'black') +
  annotate('text',x=10.5,y=15,label='C')

multiplot(Fig2A,Fig2B,Fig2C,layout=matrix(c(1,1,1,1,1,1,2,3,2,3), nrow=5, byrow=TRUE))

```

PAGEBREAK

**Figure 3**

```{r Figure 3,fig.width=5,fig.height=9}
###########
#Fig 3A

#Run the model
lm3 <- lmer(TLKB ~ TQspace + Insect + NonHelper + Sex + Tarsus + BodyMass + Helper + (1|LayYear),
            data = FlSAall,REML=FALSE)

lm3 <- standardize(lm3)

#Model averaging
lm4 <- dredge(lm3)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

Fig3A <- forestplot(table1) +
  annotate('text',x=0.75,y=2,label='A')

######################################
temp <- FlSAall
temp$Insect <- factor(temp$Insect)

ddFig3B <- ddply(temp,
                 .(Insect),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = 1.96*se(TLKB))
ddFig3B <- droplevels(ddFig3B[complete.cases(ddFig3B),])
ddFig3B$Insect <- as.numeric(paste(ddFig3B$Insect))

Fig3B <- ggplot(ddFig3B,
       aes(x = Insect,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('Annual food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  stat_smooth(data=FlSAall,aes(x=Insect,y=TLKB),method=lm,col='black',se=T) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  scale_colour_manual(values = wes_palette("Royal1")) +
  annotate('text',x=2,y=10,label='B')

multiplot(Fig3A,Fig3B)

```

PAGEBREAK

**Figure 4**

```{r Figure 4,fig.height=8,fig.width=8}
s1 <- with(subset(dd,Ageclass == 'FL'),flexsurvreg(Surv(Lifespan+1,Died)~TLKB, dist="lnorm"))

s1 <- with(chicks,flexsurvreg(Surv(Lifespan+1,Died)~TLKB+BodyMass+TLKB+Helper, dist="lnorm"))
t1 <- data.frame(s1$res)
t1 <- t1[3:5,]
t1 <- t1[order(rownames(t1)),]
rownames(t1) <- c('Body mass','Helper','Tel. length')

Fig4A <- ggplot(t1,aes(x = rownames(t1),y = est)) +
  geom_point() + 
  geom_pointrange(aes(ymin = L95.,ymax = U95.)) +
    theme_classic() +
  ylab('Estimate') + 
  xlab('') + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90,size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_abline(intercept = 0,slope = 0,lty=2) +
  annotate('text',x=0.6,y=0.4,label='A')

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig4B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

###########################################

s2 <- with(FlSAall,flexsurvreg(Surv(Lifespan+1,Died)~TLKB+Insect, dist="lnorm"))
t4 <- data.frame(s2$res)
t4 <- t4[3:4,]
t4 <- t4[order(rownames(t4)),]
rownames(t4) <- c('Food av.','Tel. length')

Fig4C <- ggplot(t4,aes(x = rownames(t4),y = est)) +
  geom_point() + 
  geom_pointrange(aes(ymin = L95.,ymax = U95.)) +
    theme_classic() +
  ylab('Estimate') + 
  xlab('') + 
  theme_classic() +
  theme(axis.text.x = element_text(angle=90,size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  geom_abline(intercept = 0,slope = 0,lty=2) +
  annotate('text',x=0.6,y=0.15,label='C')

s4 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSA)
Fig4D <- ggsurv(s4,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='D')

multiplot(Fig4A,Fig4B,Fig4C,Fig4D,layout=matrix(c(1:4), nrow=2, byrow=FALSE))

```