### 2015-09-14 Manuscript/SW_early_life_telomeres_SOM_v9.Rmd
Updated figures and manuscript - standardised all response variables for within year analyses. Added markdown docx template (myref.docx)

---
output: word_document
---

###Supplementary information for Spurgin *et al.* "Telomeres reveal complex silver spoon effects in a wild population"

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

```

**Table S1** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler chicks. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. i = intercept.

```{r Table S1}

#Run the model
lm1 <- lmer(TLKB ~ TQspace + Insect + NonHelper + Sex + BodyMass + Tarsus + Helper + (1|LayYear),
            data = chickall,REML=FALSE)

lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))
colnames(t1)[2:8] <- c('Sex', 'Body Mass','Helpers','Food av.','Non helpers','Tarsus length','TQ')

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')])
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
kable(tableS1,digits=3)

```

PAGEBREAK

**Table S2** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler fledglings and subadults. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. i = intercept.

```{r Table s2}

#Run the model
lm1 <- lmer(TLKB ~ TQspace + Insect + NonHelper + Sex + BodyMass + Tarsus + Helper + (1|LayYear),
            data = FlSAall,REML=FALSE)

lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))
colnames(t1)[2:8] <- c('Sex', 'Body Mass','Helpers','Food av.','Non helpers','Tarsus length','TQ')

tables2 <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')])
colnames(tables2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tables2) <- NULL
tables2$Model[tables2$Model == ''] <- 'Null model'
kable(tables2,digits=3)

```

PAGEBREAK

```{r Figure S1,fig.width=4,fig.height=4}

temp <- dd

ddFigS1 <- ddply(dd,
                 .(Age),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS1 <- ddFigS1[complete.cases(ddFigS1),]
ddFigS1 <- subset(ddFigS1,Age<14)

FigS1 <- ggplot(ddFigS1,
       aes(x = Age,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Age') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS1

```

**Figure S1** Cross-sectional telomere length (mean $\pm$ s.e.) in relation to age in Seychelles warblers.

PAGEBREAK

```{r Figure S2,fig.width=4,fig.height=4}

temp <- Loss
levels(temp$Ageclass) <- c('Chicks', 'Fledglings','Subadults')

ddFigS2 <- ddply(temp,
                 .(Ageclass),
                 summarise,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))

FigS2 <- ggplot(ddFigS2,
       aes(x = Ageclass,
           y = TROCmean)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS2

```

**Figure S2** Longitudinal telomere loss (mean $\pm$ s.e.) in juvenile Seychelles warblers. Telomere loss a standardised and corrected value (see methods for details) based on rates of lost between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

PAGEBREAK

```{r Figure S3,fig.width=5,fig.height=9}

###########
#Fig S3A

temp <- subset(FlSAall,Ageclass == 'FL')

#Run the model
lm3 <- lmer(TLKB ~ TQspace + Insect + Sex + Tarsus + BodyMass + (1|LayYear),
            data = temp,REML=FALSE)

lm3 <- standardize(lm3)

#Model averaging
lm4 <- dredge(lm3)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

FigS3A <- forestplot(table1) +
  annotate('text',x=0.75,y=3,label='A')

###########
#Fig S3B

temp <- subset(FlSAall,Ageclass == 'SA')

#Run the model
lm3 <- lmer(TLKB ~ TQspace + Insect + Sex + Tarsus + BodyMass + (1|LayYear),
            data = temp,REML=FALSE)

#lm3 <- standardize(lm3)

#Model averaging
lm4 <- dredge(lm3)
lmavg <- model.avg(lm4,subset = delta <=6)
myimp <- importance(lmavg) #Relative importance of each explanatory variable

#Put it all into a nice table
table1 <- imptable(lmavg,myimp)
table1 <- signif(table1,2)

#Sort out row/col names
nams <- rownames(table1)
nams[nams == 'c.Ageclass'] <- 'Age class'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.Insect'] <- 'Food av.'
nams[nams == 'z.TQspace'] <- 'TQ'
nams[nams == 'z.Helper'] <- 'Helpers'

rownames(table1) <- nams
colnames(table1) <- c('Estimate','SE','2.5%','97.5%','RI')

table1 <- data.frame(table1,myx=row.names(table1))

FigS3B <- forestplot(table1) +
  annotate('text',x=0.75,y=3,label='B')

multiplot(FigS3A,FigS3B)

```

**Figure S3** Factors affecting telomere length in Seychelles warbler fledglings **(A)** and subadults **(B)**. Points and error bars represent model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality).