### 2015-09-15 Manuscript/SW_early_life_telomeres_SOM_v9.Rmd
More manuscript updates - mainly results

---
output: word_document
---

###Supplementary information for Spurgin *et al.* "Telomeres reveal complex silver spoon effects in a wild population"

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

```

**Table S1** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler chicks. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

```{r Table S1}

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = chickall)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

```

PAGEBREAK

**Table S2** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler fledglings. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

```{r Table S2}

temp <- subset(FlSAall,Ageclass == 'FL')

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = temp)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

```

**Table S3** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler sub-adults Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

```{r Table S3}

temp <- subset(FlSAall,Ageclass == 'SA')

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = temp)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS3 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS3) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS3) <- NULL
tableS3$Model[tableS3$Model == ''] <- 'Null model'
tableS3[,3:5] <- lapply(tableS3[,3:5],as.numeric)
kable(tableS3,digits=2)

```