### 2016-11-24 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd

---
output:
  word_document:
    reference_docx: myref.docx
---
```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(include=FALSE,
               message=FALSE,
               fig.show='hide')

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')
source('Scripts/SW juvenile telomeres IPC models.r')

```

**Table 1** Telomere dynamics and age in Seychelles warbler cohorts. Linear mixed models were created with RTL as the response variable, and different measures of age, cohort ID, and cohort x age interactions as explanatory variables (see methods for details). Models are ranked by AICc, with best models at the top of the table.

```{r Table 1,include = T,echo = F,warning = F}
RTL_Age_Dredge <- dredge(RTL_Age,
                         subset = !(LogAge & FAge) & 
                           !(LogAge & Agemonths) &
                           !(LogAge & Age2) &
                           !(Agemonths & FAge) & 
                           !(Agemonths & Age2) &
                           !(FAge & Age2))

t1 <- as.matrix(RTL_Age_Dredge)

#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'Agemonths'] <- 'Age (linear)'
nams[nams == 'LogAge'] <- 'Age (log)'
nams[nams == 'FAge'] <- 'Age (factor)'
nams[nams == 'Cohort:FAge'] <- 'Age (factor)*cohort'
nams[nams == 'Agemonths:Cohort'] <- 'Age (continuous)*cohort'
nams[nams == 'Cohort:LogAge'] <- 'Age (log)*cohort'
nams[nams == 'Age2:Cohort'] <- 'Age (quadratic)*cohort'
nams[nams == 'Age2'] <- 'Age (quadratic)'

colnames(t1) <- nams

table1A <- data.frame(Model = apply(t1,1,getterms,c(2:10)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(table1A) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(table1A) <- NULL
table1A$Model[table1A$Model == ''] <- 'Null model'
table1A[,3:5] <- lapply(table1A[,3:5],as.numeric)

RTL_DeltaAge_Dredge <- dredge(RTL_DeltaAge,
                         subset = !(DeltaLogAge & DeltaAge) &
                           !(DeltaLogAge & DeltaAge2) &
                           !(DeltaAge & DeltaAge2))

t1 <- as.matrix(RTL_DeltaAge_Dredge)

#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'DeltaAge'] <- 'Delta age (linear)'
nams[nams == 'DeltaLogAge'] <- 'Delta age (log)'
nams[nams == 'Cohort:DeltaAge'] <- 'Delta age (continuous)*cohort'
nams[nams == 'Cohort:DeltaLogAge'] <- 'Delta age (log)*cohort'
nams[nams == 'Cohort:DeltaAge2'] <- 'Delta age (quadratic)*cohort'
nams[nams == 'DeltaAge2'] <- 'Delta age (quadratic)'

colnames(t1) <- nams

table1B <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(table1B) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(table1B) <- NULL
table1B$Model[table1B$Model == ''] <- 'Null model'
table1B[,3:5] <- lapply(table1B[,3:5],as.numeric)

#Sort out rounding and exlcude NAs from blank spaces. Clunky but works
table1 <- rbind(c('A',rep(' ',4)),table1A,c('--',rep(' ',4)),c('B',rep(' ',4)),table1B)
table1[,3:5] <- lapply(table1[,3:5],as.numeric)
table1[,3:5] <- round(table1[,3:5],3)
table1[,3:5] <- lapply(table1[,3:5],as.character)
table1[is.na(table1)] <- " "

kable(table1,digits = 3)
```

**Figure Legends**

**Figure 1** Telomere dynamics in relation to age in Seychelles warbler cohorts. **A** RTL and age across all individuals. Points and  thin grey lines represent individual samples and birds, respectively. The thick line and shaded area represent the fitted values and 95% confidence limits of a linear regression of RTL and log-transformed age.**B** Boxplot of variation in RTL among cohorts. For visualisation purposes a selection of cohorts with large sample sizes across a wide age range was chosen, but all cohorts were included in models. **C** RTL and age among cohorts. Lines represent fitted values from a linear regression and log-transformed age, and colours correspond to **B**. **D** RTL in relation to and $\Delta$Log age (i.e. within indiviual variation in log age).

**Figure 2** Longitudinal telomere dynamics in the Seychelles warbler. **A** Variation in RTL within individuals sampled at different time points. The dotted line represents parity, and thus points above and below the line represent increases and decreases in RTL, respectively. **B** Scaled density plots of repeated RTL measurements among individual samples, and among different samples taken from the same individual.

**Figure 3** Telomere length in relation to the social and ecological environment in the Seychelles warbler. **A** Estimates and 95% confidence intervals for all explanatory variables fitted in a linear mixed model (see methods for details). **B** RTL in relation to tarsus length and sex. **C** RTL in relation to variation in annual food availability. Lines and shaded areas represent the fitted values and 95% confidence limits from linear regressions.

```{r Figure 1}

Fig1A <- ddply(dd,
               .(BirdID),
        mutate,
        nreps = length(RTL)) %>%
  subset(nreps>0) %>%
  subset(!duplicated(paste0(BirdID,Agemonths))) %>%
  ggplot(aes(x = Agemonths,y = RTL,group = BirdID,col=factor(BirdID)))+
  geom_point(alpha = 1,col='grey')+
geom_line(col = 'darkgrey',alpha = 0.2)+
  xlab('Age (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')+
  geom_smooth(col = 'black',fill = 'black',lwd = 0.5,aes(group = NULL),method = 'lm',formula = y~log(x))+
  annotate('text',x = 0,y = 1.7,label = 'A')

c15 <- c("dodgerblue2","#E31A1C", # red
                "green4",
                "#FF7F00", # orange
                "gold1",
                "skyblue2", # lt pink
                "palegreen2",
                "#FDBF6F", # lt orange
                "gray70", "khaki2","blue1","steelblue4",
                "darkturquoise",
                "darkorange4","brown","navy",
         'yellow','green','khaki3','blue','pink','orange')

Fig1B <- transform(dd,Cohort = factor(LayYear)) %>%
  ggplot(aes(y = RTL,x = Cohort, fill = Cohort))+
  xlab('Birth year') +
  geom_boxplot(outlier.shape = NA) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
        scale_fill_manual(values = c15)+
  annotate('text',x = 1,y = 1.6,label = 'B')

Fig1C <- transform(dd,Cohort = factor(LayYear)) %>%
  ggplot(aes(y = RTL,x = Agemonths, col = Cohort,fill = Cohort))+
  xlab('Age (years)') +
  geom_smooth(method = 'lm', formula = y~log(x),se = F) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
    scale_color_manual(values = c15)+
        scale_fill_manual(values = c15)+
  annotate('text',x = 0,y = 1.3,label = 'C')

  Fig1D <- ggplot(dd3, aes(x = DeltaLogAge,y = RTL))+
  geom_point(alpha = 1,col='grey')+
  xlab(expression(Delta*'Log age (years)')) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none')+
  geom_smooth(col = 'black',method = 'lm')+
  annotate('text',x = -1.5,y = 1.7,label = 'D')

pdf('Figures/Figure 1.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout = matrix(1:4,2,2,byrow = T))
dev.off()

```

```{r Figure 2}

Fig2A <- ggplot(dd3,
       aes(x = RTL,
           y = RTL1)) +
  ylab(expression('RTL ('*italic(t)*' + 1)')) + 
  xlab(expression('RTL ('*italic(t)*')')) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = c('none'),
        legend.title = element_text(size = 12),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
    geom_point(col = 'grey')+
  geom_abline(slope = 1,intercept = 0,lty = 2) +
  annotate('text',x = 0.3,y = 1.7,label = 'A')

Fig2B <- ggplot(ddL,aes(col=Group,fill = Group,y = ..scaled..,x = DeltaRTL))+
    geom_density(alpha = 0.2)+
  theme_classic()+
  ylab ('Scaled density')+
  xlab(expression(Delta*'RTL'))+
    theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none',
        legend.title = element_blank(),
        axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black')) +
  scale_colour_manual(values = c('darkgrey','gold'))+
  scale_fill_manual(values = c('darkgrey','gold'))+
  geom_vline(xintercept = 0,lty = 2) +
  annotate('text',x = -1,y = 1,label = 'B')

pdf('Figures/Figure 2.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig2A,Fig2B)
dev.off()

```

```{r Figure 3}

CIs <- CIwald(RTL_Full2)
t1a <- data.frame(myx = row.names(CIs)[7:nrow(CIs)],
                  Estimate = fixef(RTL_Full2),
                  X2.5. = CIs[7:nrow(CIs),1],
                  X97.5. = CIs[7:nrow(CIs),2],
                  row.names = NULL,
                  stringsAsFactors = F)

nams <- t1a$myx
nams[nams == 'z.LogAge'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus length'
nams[nams == 'c.Helper'] <- 'Number helpers'
nams[nams == 'z.GroupSize'] <- 'Group size'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Insect'] <- 'Insect abundance'
nams[nams == 'z.TQ'] <- 'Territory quality'
nams[nams == 'z.Density'] <- 'Density'

t1a$myx <- nams

t1a$myx <- factor(t1a$myx,levels = t1a$myx[order(t1a$Estimate,decreasing = T)])
t1a <- subset(t1a,myx != "(Intercept)")

Fig3A <- forestplot(t1a,RI = F)+
  annotate('text',y = 0.1,x = 0.6,label = 'A')

Fig3B <- ggplot(dd,
       aes(x = Tarsus,
           y = RTL,
           col = Sex)) +
  geom_point()+
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  stat_smooth(method = 'lm') +
  scale_color_manual(values = c('gold2','red'))+
  annotate('text',x = 17,y = 1.7,label = 'B')

Fig3C <- ggplot(dd,
       aes(y = RTL,
           x=Insect)) +
    ylab('RTL') + 
  xlab('Insect abundance') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
                axis.line.x = element_line(colour = 'black'),
        axis.line.y = element_line(colour = 'black'),
        legend.position = 'none') +
  geom_boxplot(aes(group = Insect),outlier.shape = NA)+
   stat_smooth(method = 'lm',col = 'darkgrey') +
  annotate('text',x = 2.5,y = 1.6,label = 'C')

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig3A,Fig3B,Fig3C,layout = matrix(c(1,1,2,3),byrow = T,nrow = 2))
dev.off()

```